---
output:
  md_document:
    variant: markdown_github
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Rvoteview

**WARNING: This package is in flux, as is the server and data. Please limit the number of roll calls you request at a time and be prepared for queries to break or behave strangely. See usage below.**

This is a package that enables you to query the Voteview database for roll calls and work with data frames or a `pscl` `rollcall` object.

To install this package, ensure you have `devtools` installed. If you do not, run `install.packages("devtools")` before doing the following:
```R
devtools::install_github("JeffreyBLewis/Rvoteview")
```
  
## Using Rvoteview

To use `Rvoteview`, you generally want to query the database to get a list of vote ids and then use those to return the individual votes. We query the database with a search term like so:
```{r search}
library(Rvoteview)
  
res <- voteview.search("Iraq")
names(res)
  
## I will drop description since it is a very long field
head(res[, -2])
```
Using `res$id` we can get a `rollcall` object (from the [`pscl` package](https://cran.r-project.org/web/packages/pscl/index.html)) that contains the full set of votes and data for each roll call. Eventually we will either develop additional methods for the `pscl` `rollcall` object.
```{r download}
## Get a rollcall object using the ids, please limit to a few ids for now!
rc <- voteview.download(res$id[1:10])

## Now this object can be used in many 'pscl' methods
summary(rc)
```

Please see the help files for each function after you install the package to see a little more about how they work.
