---
output:
  md_document:
    variant: markdown_github
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Rvoteview

**WARNING: This package is under construction. The server and database are currently being revamped. Not all features work, so please limit usage to the simple kind of examples you see below. Also, please limit the number of roll calls you request at a time and be prepared for queries to break or behave strangely.**

This is a package that enables you to query the Voteview database for roll calls and work with data frames or a `pscl` `rollcall` object.

To install this package, ensure you have `devtools` installed. If you do not, run `install.packages("devtools")` before doing the following:
```R
devtools::install_github("JeffreyBLewis/Rvoteview")
```

For more thorough documentation, see the help files and the vignette included with the package:
```R
vignette("voteview-vignette")
```

## Using Rvoteview

To use `Rvoteview`, you generally want to query the database to get a list of vote ids and then use those to return the individual votes. We query the database with a search term and some parameters to constrain the search. For now we take a string that will seach for a bill that has ANY of the words in the string in ANY text field. Additional boolean and targeted searches have been implemented (and will be explained in the vignette and documentation soon).
```{r search}
library(Rvoteview)
  
res <- voteview_search("Iraq")
names(res)
  
## I will drop description since it is a very long field
head(res[, -1])
```
Using `res$id` we can get a `rollcall` object (from the [`pscl` package](https://cran.r-project.org/web/packages/pscl/index.html)) that contains the full set of votes and data for each roll call. 

```{r download, results='hide'}
## Get a rollcall object using the ids, please limit to a few ids for now!
rc <- voteview_download(res$id[1:10])
```
```{r pscl-sum}
## Now this object can be used in many 'pscl' methods
summary(rc)
```

We have also begun extending the `pscl` `rollcall` object by allowing you to do a full outer join on two `rollcall` objects that were built using this package. For example:

```{r download2, results='hide'}
rc2 <- voteview_download(res$id[c(5, 100)])
```
```{r pscl-add}
## Now we can merge this with 'rc'
rcAll <- rc %+% rc2

rc$m
rc2$m
rcAll$m # rc$m + rc2$m - 1 because of the one vote overlap
```

You can also search by start and end date, congress, and chamber. Please see the help files for each function after you install the package to see a little more about how they work.

```{r search-options}
## Voteview search with options
res <- voteview_search("Iraq", chamber = "House", congress = 110,
                       startdate = 2008, enddate = "2008-04-20")
head(res[, -1])

res <- voteview_search("Iraq", congress = 109:112)
head(res[, -1])
```

We can also build complex queries manually (note session will be replaced with congress as soon as the database is updated).

```{r search-adv}
## Voteview search with options
res <- voteview_search("((alltext:'estate tax' OR 'death tax') AND session:[100 to 114]) OR (alltext:'property tax' AND session:[95 to 100])")
head(res[, -1])
tail(res[, -1])
```

